---
title: "2019-08-25-hatch-ph-multifactorial-merged"
author: "Michael J. Braus"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    number_sections: false
---

\fontsize{12}{14}
\selectfont{}

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path = 'output/', echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev="png", dpi=300)
```

# Preliminary Look


## All Data: 
[...]

```{r}
library(ggplot2)
```

```{r}
dat <- read.csv(file = "multifactorial/2019-08-25-hatch-ph-multifactorial-merged.csv", header = T)
dat$co2.pH.diff <- dat$Lab.CO2.MicropH-dat$High.CO2.MicropH
dat$co2.exp.pH.diff <- 10^-(dat$Lab.CO2.MicropH)-10^-(dat$High.CO2.MicropH)
str(dat)
```

```{r}
qplot(data = dat, x = Lab.CO2.MicropH, y = High.CO2.MicropH, color = Water.Soil.Ratio) + scale_x_continuous(limits = c(4.8, 8.1)) + scale_y_continuous(limits = c(4.8, 8.1)) + theme_bw() + geom_smooth(method = "lm") + geom_abline(slope = 1, intercept = 0, color = "black")
```

```{r}
qplot(data = dat, x = Lab.CO2.MicropH, y = High.CO2.MicropH, color = Station.Name)
```

```{r}
qplot(data = dat, x = 10^-Lab.CO2.MicropH, y = 10^-High.CO2.MicropH, color = Water.Soil.Ratio) + scale_x_continuous(limits = c(10^-8.1, 10^-4.8)) + scale_y_continuous(limits = c(10^-8.1, 10^-4.8)) + theme_bw() + geom_smooth(method = "lm") + geom_abline(slope = 1, intercept = 0, color = "black")
```

```{r}
qplot(data = dat, x = 10^-Lab.CO2.MicropH, y = 10^-High.CO2.MicropH, color = Station.Name)
```

```{r}
qplot(dat$co2.pH.diff, geom="histogram", color = dat$Station.Name, alpha = 0.4, binwidth = 0.05) 
```

```{r}
qplot(100*(dat$co2.exp.pH.diff)/(10^-dat$Lab.CO2.MicropH), geom="histogram", color = dat$Station.Name, alpha = 0.4, binwidth = 3) 
```

```{r}
t.test(dat$Lab.CO2.MicropH, dat$High.CO2.MicropH, paired = TRUE, alternative = "two.sided")
```

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = Lab.CO2.MicropH, color = Station.Name)
```

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = 10^-Lab.CO2.MicropH, color = Water.Soil.Ratio)
```

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = 10^-Lab.CO2.MicropH, color = Station.Name)
```

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = 10^-Lab.CO2.MicropH, color = Texture.Name)
```

```{r}
qplot(data = dat, x = Perc.Clay, y = 10^-Lab.CO2.MicropH, color = Station.Name)
```

## Replication of Jackson c.1950: 

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = Lab.CO2.MicropH, color = Sample.ID) + geom_line() + theme(legend.position="none")
```

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = 10^-Lab.CO2.MicropH, color = Sample.ID) + geom_line() + theme(legend.position="none")
```

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = High.CO2.MicropH, color = Sample.ID) + geom_line() + theme(legend.position="none")
```

```{r}
qplot(data = dat, x = Real.Water.Soil.ratio, y = 10^-High.CO2.MicropH, color = Sample.ID) + geom_line() + theme(legend.position="none")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = Lab.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = 10^-Lab.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = High.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = 10^-High.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = High.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none") + facet_wrap(~Study)
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = 10^-High.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none") + facet_wrap(~Study)
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = Lab.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none") + facet_wrap(~Station.Name, scales="free")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = 10^-Lab.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none") + facet_wrap(~Station.Name, scales="free")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = High.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none") + facet_wrap(~Station.Name, scales="free")
```

```{r}
p <- ggplot(dat, aes(x = Real.Water.Soil.ratio, y = 10^-High.CO2.MicropH, color = Sample.ID))
p + geom_point() + geom_smooth(method = "glm", se=F) + theme(legend.position="none") + facet_wrap(~Station.Name, scales="free")
```

## Just Spooner: 

```{r}
dat.Spooner <- subset(dat, Station.Name=="Spooner")
t.test(dat.Spooner$Lab.CO2.MicropH, dat.Spooner$High.CO2.MicropH, paired = TRUE, alternative = "two.sided")
t.test(10^-dat.Spooner$Lab.CO2.MicropH, 10^-dat.Spooner$High.CO2.MicropH, paired = TRUE, alternative = "two.sided")
```

```{r}
qplot(data = dat.Spooner, x = Lab.CO2.MicropH, y = High.CO2.MicropH, color = Water.Soil.Ratio)
```

```{r}
qplot(data = dat.Spooner, x = 10^-Lab.CO2.MicropH, y = 10^-High.CO2.MicropH, color = Water.Soil.Ratio)
```

```{r}
qplot(dat.Spooner$co2.pH.diff, geom="histogram", binwidth = 0.05) 
```

```{r}
qplot(100*(dat.Spooner$co2.exp.pH.diff)/(10^-dat.Spooner$Lab.CO2.MicropH), geom="histogram", binwidth = 4) 
```

## Just Wisconsin: 

```{r}
dat.Wisconsin <- subset(dat, Station.Name!="Spooner")
t.test(dat.Wisconsin$Lab.CO2.MicropH, dat.Wisconsin$High.CO2.MicropH, paired = TRUE, alternative = "two.sided")
t.test(10^-dat.Wisconsin$Lab.CO2.MicropH, 10^-dat.Wisconsin$High.CO2.MicropH, paired = TRUE, alternative = "two.sided")
```

```{r}
qplot(data = dat.Wisconsin, x = Lab.CO2.MicropH, y = High.CO2.MicropH, color = Water.Soil.Ratio)
```

```{r}
qplot(data = dat.Wisconsin, x = Lab.CO2.MicropH, y = High.CO2.MicropH, color = Station.Name)
```

```{r}
qplot(data = dat.Wisconsin, x = 10^-Lab.CO2.MicropH, y = 10^-High.CO2.MicropH, color = Water.Soil.Ratio)
```

```{r}
qplot(dat.Wisconsin$co2.pH.diff, geom="histogram", color = dat.Wisconsin$Station.Name, alpha = 0.4, binwidth = 0.05) 
```

```{r}
qplot(100*(10^-dat.Wisconsin$co2.exp.pH.diff)/(10^-dat.Wisconsin$Lab.CO2.MicropH), geom="histogram", color = dat.Wisconsin$Station.Name, alpha = 0.4, binwidth = 3) 
```


# Past Preliminary Analysis

We can split the data according to sample sets: "CrossWisc" and "Spooner"

```{r}
dat.CrossWisc <- subset(dat, Sample.Number <= 65)
str(dat.CrossWisc)
```

```{r}
dat.Spooner <- subset(dat, Sample.Number >= 185)
str(dat.Spooner)
```

We can plot of the dry weight and real water soil:soil ratios of each group of data (1:1, 1:2, 1:3, & 1:4 of "CrossWisc" and "Spooner" soil sample sets): 

## CrossWisc

### 1:1

```{r}
dat.CrossWisc.1to1 <- subset(dat.CrossWisc, Water.Soil.Ratio=="OneToOne")
plot(dat.CrossWisc.1to1$Real.Water.Soil.ratio, dat.CrossWisc.1to1$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```


### 1:2

```{r}
dat.CrossWisc.1to2 <- subset(dat.CrossWisc, Water.Soil.Ratio=="OneToTwo")
plot(dat.CrossWisc.1to2$Real.Water.Soil.ratio, dat.CrossWisc.1to2$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```


### 1:3

```{r}
dat.CrossWisc.1to3 <- subset(dat.CrossWisc, Water.Soil.Ratio=="OneToThree")
plot(dat.CrossWisc.1to3$Real.Water.Soil.ratio, dat.CrossWisc.1to3$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```


### 1:4

```{r}
dat.CrossWisc.1to4 <- subset(dat.CrossWisc, Water.Soil.Ratio=="OneToFour")
plot(dat.CrossWisc.1to4$Real.Water.Soil.ratio, dat.CrossWisc.1to4$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```

They all have a similar pattern. 

## Spooner

### 1:1

```{r}
dat.Spooner.1to1 <- subset(dat.Spooner, Water.Soil.Ratio=="OneToOne")
plot(dat.Spooner.1to1$Real.Water.Soil.ratio, dat.Spooner.1to1$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```


### 1:2

```{r}
dat.Spooner.1to2 <- subset(dat.Spooner, Water.Soil.Ratio=="OneToTwo")
plot(dat.Spooner.1to2$Real.Water.Soil.ratio, dat.Spooner.1to2$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```


### 1:3

```{r}
dat.Spooner.1to3 <- subset(dat.Spooner, Water.Soil.Ratio=="OneToThree")
plot(dat.Spooner.1to3$Real.Water.Soil.ratio, dat.Spooner.1to3$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```


### 1:4

```{r}
dat.Spooner.1to4 <- subset(dat.Spooner, Water.Soil.Ratio=="OneToFour")
plot(dat.Spooner.1to4$Real.Water.Soil.ratio, dat.Spooner.1to4$Dry.Soil.g, xlim = c(0,2), ylim = c(0,2))
```

The Spooner sample set has a similar pattern, but it's more grouped than the CrossWisc sample set. This makes sense because their textures are nearly identical. 


# Plots with 16S Amplicon Analysis: 

```{r, results="hide" }
# Intall:
# BiocManager::install("phyloseq")
# Load libraries
library(phyloseq)
library(dada2)
library(ggplot2)
library(RColorBrewer)
library(vegan)
```

```{r, results="hide" }
# Import sequence processing table and check it out
# How many sequences were retained at each step?
track = readRDS("reads/track.rds")
track
```

```{r, results="hide"}
# Import final otu table
OTUs = readRDS("reads/OTUtab.nochim.rds")
# Change rows and columns so taxa are rows
OTUs = t(OTUs)
# Import as phyloseq object table
otutab = otu_table(OTUs, taxa_are_rows=TRUE)
head(otutab)
# Look at lowest-abundance samples
head(sort(sample_sums(otutab)))
name = row.names(otutab)[1]
name
hist(nchar(row.names(otutab)))
```

```{r, results="hide"}
colnames(otutab)
# The sample names have suffixes appended to them - we just want sample IDs
# Get names
names = colnames(otutab)
# Remove last 17 characters
newnames = gsub('.{17}$', '', names)
# Check out names
newnames

# Assign new names to samples
colnames(otutab) = newnames

# Same thing if we wanted to simplify OTU names to IDs
# Instead of sequences (although that's useful)
otunames=row.names(otutab)
length(otunames)
newotunames = paste("OTU",rep(1:length(otunames)),sep="")
head(newotunames)
row.names(otutab) = newotunames
head(otutab)

# Cut off the sampling ID to match metadata
for (i in 1:length(colnames(otutab))){
  newname = strsplit(colnames(otutab[i]),"_")[[i]][1]
  colnames(otutab)[i]=newname
}

# Import sample data
samdat = read.csv("hatch-metadata-spooner-copy.csv", header=T)
# samdat$Sample.ID = as.character(samdat$Sample.ID)
# samdat$Sample.ID[1:9] = paste("00",samdat$Sample.ID[1:9],sep="")
# samdat$Sample.ID[10:length(samdat$Sample.ID)] = paste("0",samdat$Sample.ID[10:length(samdat$Sample.ID)],sep="")

# Check we have all the same sample names now
samdat$MARS.Plot.exp.pH <- as.numeric(10^(-samdat$MARS.Plot.pH))
samdat$Sample.ID == colnames(otutab)[66:125]
samdat = sample_data(samdat)

row.names(samdat) = samdat$Sample.ID

# Create phyloseq object
ps = phyloseq(otu_table=otutab,sample_data=samdat)
```

```{r, results="hide"}
ps.melt = psmelt(ps)
ps.melt

head(ps.melt)

hist(sample_sums(ps))

# Relative abundances
ps.norm = transform_sample_counts(ps,function(x) x/sum(x))
head(otu_table(ps.norm))
sample_sums(ps.norm)
# Looks good

colnames(sample_data(ps.norm))
```

```{r PCoA_bray, fig.height=5, fig.width=6}
row.names(samdat)
ps.ord.pcoa = ordinate(ps, "PCoA", "bray")
p = plot_ordination(ps, ps.ord.pcoa, type="samples",label="Names")
p
```

```{r NMDS_bray_SoilpH, fig.height=5, fig.width=6}
ps.ord.nmds = ordinate(ps.norm, "NMDS", "bray",trymax=1000,k=2)
p = plot_ordination(ps.norm, ps.ord.nmds, type="samples")
palette = brewer.pal(8, "Spectral")
p = p + aes(colour=MARS.Plot.pH) + geom_point(size=3) + scale_colour_gradientn(colors=palette)
p
```

```{r richness_pH}
p <- plot_richness(ps, x="MARS.Plot.pH", measures=c("Observed","Shannon"))
p + geom_point(size=5, alpha=0.7)
```

```{r permANOVA_1 }
df = as(sample_data(ps.norm), "data.frame")
d = phyloseq::distance(ps.norm, "bray") 
ps.adonis = adonis(d ~ MARS.Plot.exp.pH+MARS.Plot.pH+MARS.Plot.OM.perc+MARS.Plot.P.ppm+MARS.Plot.K.ppm, df)
ps.adonis
```


# Analysis from Thea: 

```{r}
library(phyloseq)
library(dada2)
library(ggplot2)
library(RColorBrewer)
# Load libraries

# Import sequence processing table and check it out
# How many sequences were retained at each step?
track = readRDS("track.rds")
track

# Import final otu table
OTUs = readRDS("OTUtab.nochim.rds")
# Change rows and columns so taxa are rows
OTUs = t(OTUs)
# Import as phyloseq object table
otutab = otu_table(OTUs, taxa_are_rows=TRUE)
head(otutab)
# Look at lowest-abundance samples
head(sort(sample_sums(otutab)))

name = row.names(otutab)[1]
name
hist(nchar(row.names(otutab)))

colnames(otutab)
# The sample names have suffixes appended to them - we just want sample IDs
# Get names
names = colnames(otutab)
# Remove last 17 characters
newnames = gsub('.{17}$', '', names)
# Check out names
newnames

# Assign new names to samples
colnames(otutab) = newnames

# Same thing if we wanted to simplify OTU names to IDs
# Instead of sequences (although that's useful)
otunames=row.names(otutab)
length(otunames)
newotunames = paste("OTU",rep(1:length(otunames)),sep="")
head(newotunames)
row.names(otutab) = newotunames
head(otutab)

# Cut off the sampling ID to match metadata
for (i in 1:length(colnames(otutab))){
  newname = strsplit(colnames(otutab[i]),"_")[[i]][1]
  colnames(otutab)[i]=newname
}

# Import sample data
samdat = read.csv("Hatch-Metadata.csv")
samdat$Sample.ID = as.character(samdat$Sample.ID)
samdat$Sample.ID[1:9] = paste("00",samdat$Sample.ID[1:9],sep="")
samdat$Sample.ID[10:length(samdat$Sample.ID)] = paste("0",samdat$Sample.ID[10:length(samdat$Sample.ID)],sep="")

# Check we have all the same sample names now
samdat$Sample.ID == colnames(otutab)[1:65]
samdat = sample_data(samdat)

row.names(samdat) = samdat$Sample.ID

# Create phyloseq object
ps = phyloseq(otu_table=otutab,sample_data=samdat)

ps.melt = psmelt(ps)
ps.melt

head(ps.melt)

hist(sample_sums(ps))

# Relative abundances
ps.norm = transform_sample_counts(ps,function(x) x/sum(x))
head(otu_table(ps.norm))
sample_sums(ps.norm)
# Looks good

ps.norm.noC = prune_samples(sample_names(ps.norm) != "049-W3-Compost",ps.norm)
ps.norm.noC = prune_samples(sample_names(ps.norm.nocompost) != "004-K2-Muck",ps.norm.nocompost)
ps.norm.noC

colnames(sample_data(ps.norm))

ps.ord.nmds = ordinate(ps.norm.noC, "NMDS", "bray",trymax=1000,k=2)
p = plot_ordination(ps.norm.noC, ps.ord.nmds, type="samples")
palette = brewer.pal(8, "Spectral")
p = p + aes(colour=Soil.pH) + geom_point(size=3) + scale_colour_gradientn(colors=palette)
p

row.names(samdat)
ps.ord.pcoa = ordinate(ps, "PCoA", "bray")
p = plot_ordination(ps, ps.ord.pcoa, type="samples",label="Names")
p
```