---
title: "2020-03-12-hatch-final-data-merged"
author: "Michael J. Braus"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    number_sections: false
---

\fontsize{12}{14}
\selectfont{}

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path = 'output-rmd/', echo = TRUE, warning = FALSE, message = FALSE, cache = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_chunk$set(dev="png", dpi=300)
```

# METROLOGY PAPER

Let's load some packages and the data set. 

```{r prep-metrology}
library(ggplot2)
library(leaps)
library(cowplot)
dat <- read.csv(file = "source/metrology-compiled/2020-03-12-hatch-data-merged.csv", header = T)
str(dat)
```

Let's process the data a little by making exponentiated pH values, making diffs, etc. 

```{r process-dat }
dat$Study <- as.character(dat$Study)
dat$Mean.Depth <- (dat$Upper.Depth.cm+dat$Lower.Depth.cm)/2
```

```{r jackson-plot-ph, fig.height=8, fig.width=6}
ggplot(subset(dat, Study!=""&Real.Water.Soil.ratio<=1.3), aes(Real.Water.Soil.ratio, Lab.CO2.pH, color=Sample.ID, shape=Horizon)) + 
  geom_point(size=1) + geom_line(size=0.3) + theme_bw() + theme(legend.position="none") +
  facet_wrap(~Study, ncol=1) + labs(x="Water/Soil by Mass", y="Soil pH") + 
  scale_y_continuous(breaks = seq(4, 8, by = 0.25))
```

```{r jackson-plot-ph-wisc, fig.height=8, fig.width=8}
jack.wisc <- ggplot(subset(dat, Study!=""&Study!="Spooner"&Real.Water.Soil.ratio<=1.3), aes(Real.Water.Soil.ratio, Lab.CO2.pH, color=Sample.ID, shape=Horizon)) + 
  geom_point(size=3) + geom_point(colour = "grey90", size = 1.5) + geom_line(size=0.3) + theme_bw() + theme(legend.position="none") +
  facet_wrap(~Station.Name, ncol=3) + labs(x="Water/Soil by Mass", y="Soil pH") + 
  scale_y_continuous(breaks = seq(4, 8, by = 0.25))+
  scale_x_continuous(breaks = seq(0, 1.25, by = 0.2))
jack.wisc
```

```{r jackson-plot-ph-spooner, fig.height=5, fig.width=6}
jack.spoon <- ggplot(subset(dat, Study!=""&Study!="Wisconsin"&Real.Water.Soil.ratio<=1.3), aes(Real.Water.Soil.ratio, Lab.CO2.pH, color=Sample.ID, shape=Horizon)) + 
  geom_point(size=3) + geom_point(colour = "grey90", size = 1.5) + geom_line(size=0.3) + theme_bw() + theme(legend.position="none") +
  facet_wrap(~Station.Name, ncol=3) + labs(x="Water/Soil by Mass", y="Soil pH") + 
  scale_y_continuous(breaks = seq(4, 8, by = 0.25))+
  scale_x_continuous(breaks = seq(0, 1.25, by = 0.2))
jack.spoon
```

```{r jackson-plot-activity-wisc, fig.height=8, fig.width=8}
ggplot(subset(dat, Study!=""&Study!="Spooner"&Real.Water.Soil.ratio<=1.3), aes(Real.Water.Soil.ratio, 10^-Lab.CO2.pH, color=Sample.ID, shape=Horizon)) + 
  geom_point(size=2) + geom_line(size=0.3) + theme_bw() + theme(legend.position="none") +
  facet_wrap(~Station.Name, ncol=3) + labs(x="Water/Soil by Mass", y="Soil a(H+)")
```

```{r jackson-plot-activity-spooner, fig.height=5, fig.width=6}
ggplot(subset(dat, Study!=""&Study!="Wisconsin"&Real.Water.Soil.ratio<=1.3), aes(Real.Water.Soil.ratio, 10^-Lab.CO2.pH, color=Sample.ID, shape=Horizon)) + 
  geom_point(size=2) + geom_line(size=0.3) + theme_bw() + theme(legend.position="none") +
  facet_wrap(~Station.Name, ncol=3) + labs(x="Water/Soil by Mass", y="Soil a(H+)")
```

```{r jackson-plot-wiscons-spooner, fig.width=12, fig.height=6}
plot_grid(jack.spoon, jack.wisc)
```

## Wisconsin Soil pH ~ Nutrient Profiles

This is the BIC for the Wisconsin soils, using the available "metadata", which means soil chemical data. 

```{r subset-wisconsin }
dat.wisc <- subset(dat, Study=="Wisconsin")
```

### one2one

```{r subset-wisc.one2one }
dat.wisc.1.to.1 <- subset(dat.wisc, Water.Soil.Ratio=="1-to-1")
str(dat.wisc.1.to.1)
```

```{r bic.wisc.one2one.labco2 }
bic.wisc.one2one.labco2 <- regsubsets(Lab.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2, data = dat.wisc.1.to.1)
plot(bic.wisc.one2one.labco2, scale='bic', main="Wisconsin Soils, 1-to-1 Water-to-Soil, ~400 ppm CO2")
```

```{r bic.wisc.one2one.highco2 }
bic.wisc.one2one.highco2 <- regsubsets(High.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2,data = dat.wisc.1.to.1)
plot(bic.wisc.one2one.highco2, scale='bic', main="Wisconsin Soils, 1-to-1 Water-to-Soil, ~2% CO2")
```

### one2two

```{r subset-wisc.one2two}
dat.wisc.1.to.2 <- subset(dat.wisc, Water.Soil.Ratio=="1-to-2")
str(dat.wisc.1.to.2)
```

```{r bic.wisc.one2two.labco2}
bic.wisc.one2two.labco2 <- regsubsets(Lab.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2,data = dat.wisc.1.to.2)
plot(bic.wisc.one2two.labco2, scale='bic', main="Wisconsin Soils, 1-to-2 Water-to-Soil, ~400 ppm CO2")
```

```{r bic.wisc.one2two.highco2}
bic.wisc.one2two.highco2 <- regsubsets(High.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2, data = dat.wisc.1.to.2)
plot(bic.wisc.one2two.highco2, scale='bic', main="Wisconsin Soils, 1-to-2 Water-to-Soil, ~2% CO2")
```

### one2three

```{r subset-wisc.one2three}
dat.wisc.1.to.3 <- subset(dat.wisc, Water.Soil.Ratio=="1-to-3")
str(dat.wisc.1.to.3)
```

```{r bic.wisc.one2three.labco2}
bic.wisc.one2three.labco2 <- regsubsets(Lab.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2,data = dat.wisc.1.to.3)
plot(bic.wisc.one2three.labco2, scale='bic', main="Wisconsin Soils, 1-to-3 Water-to-Soil, ~400 ppm CO2")
```

```{r bic.wisc.one2three.highco2}
bic.wisc.one2three.labco2 <- regsubsets(High.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2,data = dat.wisc.1.to.3)
plot(bic.wisc.one2three.labco2, scale='bic', main="Wisconsin Soils, 1-to-3 Water-to-Soil, ~2% CO2")
```

### one2four

```{r subset-wisc.one2four}
dat.wisc.1.to.4 <- subset(dat.wisc, Water.Soil.Ratio=="1-to-4")
str(dat.wisc.1.to.4)
```

```{r bic.wisc.one2four.labco2}
bic.wisc.one2four.labco2 <- regsubsets(Lab.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2,data = dat.wisc.1.to.4)
plot(bic.wisc.one2four.labco2, scale='bic', main="Wisconsin Soils, 1-to-4 Water-to-Soil, ~400 ppm CO2")
```

```{r bic.wisc.one2four.highco2}
bic.wisc.one2four.highco2 <- regsubsets(High.CO2.pH ~ Perc.Sand + Perc.Silt + Perc.Clay + OM.perc + Total.N.perc + Total.Org.C.perc + Bray.1.P.ppm + K.ppm + Ca.ppm + Mg.ppm + Est.Ac.Hplus + CEC.sum + Mean.Depth, nbest = 2,data = dat.wisc.1.to.4)
plot(bic.wisc.one2four.highco2, scale='bic', main="Wisconsin Soils, 1-to-4 Water-to-Soil, ~2% CO2")
```

## Soil pH Values

### Wisconsin 

```{r wisc-stand-sim-soil-ph}
q1 <- qplot(data = dat.wisc, x = Lab.CO2.pH, y = High.CO2.pH, color = Water.Soil.Ratio, shape = Water.Soil.Ratio) + geom_point(size=2) +
  theme_bw() + geom_smooth(method = "glm") + 
  geom_abline(slope = 1, intercept = 0, color = "black") + 
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "pH of Extract at 400 ppm carbon dioxide", 
       y = "pH of Extract at 2% carbon dioxide")  + 
  ggtitle("Wisconsin")
q1
```

```{r wisc-stand-sim-soil-hplus}
q2 <- qplot(data = dat.wisc, x = 10^-Lab.CO2.pH, y = 10^-High.CO2.pH, color = Water.Soil.Ratio, shape = Water.Soil.Ratio) + geom_point(size=2) + scale_x_continuous(limits = c(10^-8.1, 10^-4.8)) + scale_y_continuous(limits = c(10^-8.1, 10^-4.8)) + theme_bw() + geom_smooth(method = "glm") + geom_abline(slope = 1, intercept = 0, color = "black")+
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "a(H+)[M] of Extract at 400 ppm carbon dioxide", 
       y = "a(H+)[M] of Extract at 2% carbon dioxide") + 
  ggtitle("Wisconsin")
q2
```

### Spooner

```{r subset-spooner}
dat.spooner <- subset(dat, Study=="Spooner")
```

```{r spooner-stand-sim-soil-ph}
q3 <- qplot(data = dat.spooner, x = Lab.CO2.pH, y = High.CO2.pH, color = Water.Soil.Ratio, shape = Water.Soil.Ratio) + geom_point(size=2) +
  theme_bw() + geom_smooth(method = "glm") + 
  geom_abline(slope = 1, intercept = 0, color = "black") + 
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "pH of Extract at 400 ppm carbon dioxide", 
       y = "pH of Extract at 2% carbon dioxide")  + 
  ggtitle("Spooner")
q3
```

```{r spooner-stand-sim-soil-hplus}
q4 <- qplot(data = dat.spooner, x = 10^-Lab.CO2.pH, y = 10^-High.CO2.pH, color = Water.Soil.Ratio, shape = Water.Soil.Ratio) + geom_point(size=2) + scale_x_continuous(limits = c(10^-8.1, 10^-4.8)) + scale_y_continuous(limits = c(10^-8.1, 10^-4.8)) + theme_bw() + geom_smooth(method = "glm") + geom_abline(slope = 1, intercept = 0, color = "black")+
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "a(H+)[M] of Extract at 400 ppm carbon dioxide", 
       y = "a(H+)[M] of Extract at 2% carbon dioxide") + 
  ggtitle("Spooner")
q4
```

### Grid of Multifactorial Plots

```{r multifactoral-grid-ph, fig.width=10, fig.height=7}
plot_grid(q1, q2, q3, q4)
```


*BLACK AND WHITE*

```{r wisc-stand-sim-soil-ph-bw}
q1.bw <- qplot(data = dat.wisc, x = Lab.CO2.pH, y = High.CO2.pH, shape = Water.Soil.Ratio) + geom_point(size=2) +
  theme_bw() + geom_smooth(method = "glm") + 
  geom_abline(slope = 1, intercept = 0, color = "black") + 
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "pH of Extract at 400 ppm carbon dioxide", 
       y = "pH of Extract at 2% carbon dioxide")  + 
  ggtitle("Wisconsin")
q1.bw
```

```{r wisc-stand-sim-soil-hplus-bw}
q2.bw <- qplot(data = dat.wisc, x = 10^-Lab.CO2.pH, y = 10^-High.CO2.pH, shape = Water.Soil.Ratio) + geom_point(size=2) + scale_x_continuous(limits = c(10^-8.1, 10^-4.8)) + scale_y_continuous(limits = c(10^-8.1, 10^-4.8)) + theme_bw() + geom_smooth(method = "glm") + geom_abline(slope = 1, intercept = 0, color = "black")+
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "a(H+)[M] of Extract at 400 ppm carbon dioxide", 
       y = "a(H+)[M] of Extract at 2% carbon dioxide") + 
  ggtitle("Wisconsin")
q2.bw
```

### Spooner

```{r spooner-stand-sim-soil-ph-bw}
q3.bw <- qplot(data = dat.spooner, x = Lab.CO2.pH, y = High.CO2.pH, shape = Water.Soil.Ratio) + geom_point(size=2) +
  theme_bw() + geom_smooth(method = "glm") + 
  geom_abline(slope = 1, intercept = 0, color = "black") + 
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "pH of Extract at 400 ppm carbon dioxide", 
       y = "pH of Extract at 2% carbon dioxide")  + 
  ggtitle("Spooner")
q3.bw
```

```{r spooner-stand-sim-soil-hplus-bw}
q4.bw <- qplot(data = dat.spooner, x = 10^-Lab.CO2.pH, y = 10^-High.CO2.pH, shape = Water.Soil.Ratio) + geom_point(size=2) + scale_x_continuous(limits = c(10^-8.1, 10^-4.8)) + scale_y_continuous(limits = c(10^-8.1, 10^-4.8)) + theme_bw() + geom_smooth(method = "glm") + geom_abline(slope = 1, intercept = 0, color = "black")+
  labs(color='Water-to-Soil \n Ratio by Mass', 
       x = "a(H+)[M] of Extract at 400 ppm carbon dioxide", 
       y = "a(H+)[M] of Extract at 2% carbon dioxide") + 
  ggtitle("Spooner")
q4.bw
```

### Grid of Multifactorial Plots

```{r multifactoral-grid-ph-bw, fig.width=10, fig.height=7}
plot_grid(q1.bw, q2.bw, q3.bw, q4.bw)
```




# MICROBIAL PAPER

## Biom + Metadata

Let's load all these packages. 

```{r prep-microbial, results="hide" }
# Intall:
# BiocManager::install("phyloseq")
# Load libraries
library(phyloseq)
#library(dada2)
library(ggplot2)
library(RColorBrewer)
library(vegan)
```

Here we'll load in the microbial data. 

```{r load-biom}
# Import sequence processing table and check it out
# How many sequences were retained at each step?
track = readRDS("reads/track.rds")
track
# Import final otu table
OTUs = readRDS("reads/OTUtab.nochim.rds")
# Change rows and columns so taxa are rows
OTUs = t(OTUs)
# Import as phyloseq object table
otutab = otu_table(OTUs, taxa_are_rows=TRUE)
head(otutab)
# Look at lowest-abundance samples
head(sort(sample_sums(otutab)))
name = row.names(otutab)[1]
name
hist(nchar(row.names(otutab)))
colnames(otutab)
# The sample names have suffixes appended to them - we just want sample IDs
# Get names
names = colnames(otutab)
# Remove last 17 characters
newnames = gsub('.{17}$', '', names)
# Check out names
newnames

# Assign new names to samples
colnames(otutab) = newnames

# Same thing if we wanted to simplify OTU names to IDs
# Instead of sequences (although that's useful)
otunames=row.names(otutab)
length(otunames)
newotunames = paste("OTU",rep(1:length(otunames)),sep="")
head(newotunames)
row.names(otutab) = newotunames
head(otutab)

# Cut off the sampling ID to match metadata
for (i in 1:length(colnames(otutab))){
  newname = strsplit(colnames(otutab[i]),"_")[[i]][1]
  colnames(otutab)[i]=newname
}

# Import sample data
samdat = read.csv("source/metrology-compiled/2020-03-12-hatch-data-merged-as-metadata.csv", header=T)
# samdat$Sample.ID = as.character(samdat$Sample.ID)
# samdat$Sample.ID[1:9] = paste("00",samdat$Sample.ID[1:9],sep="")
# samdat$Sample.ID[10:length(samdat$Sample.ID)] = paste("0",samdat$Sample.ID[10:length(samdat$Sample.ID)],sep="")

# Check we have all the same sample names now
samdat$Sample.ID[1:65] == colnames(otutab)[1:65]
samdat$Sample.ID[66:125] == colnames(otutab)[66:125]
samdat = sample_data(samdat)

# Exponentiate all pH values
samdat$DNA.Extr.Hplus.After.C1 <- 10^-samdat$DNA.Extr.pH.After.C1
samdat$DNA.Extr.Hplus.After.C2 <- 10^-samdat$DNA.Extr.pH.After.C2
samdat$Lab.CO2.Hplus.one2one <- 10^-samdat$Lab.CO2.pH.one2one
samdat$Lab.CO2.Hplus.one2two <- 10^-samdat$Lab.CO2.pH.one2two
samdat$Lab.CO2.Hplus.one2three <- 10^-samdat$Lab.CO2.pH.one2three
samdat$Lab.CO2.Hplus.one2four <- 10^-samdat$Lab.CO2.pH.one2four
samdat$High.CO2.Hplus.one2one <- 10^-samdat$High.CO2.pH.one2one
samdat$High.CO2.Hplus.one2two <- 10^-samdat$High.CO2.pH.one2two
samdat$High.CO2.Hplus.one2three <- 10^-samdat$High.CO2.pH.one2three
samdat$High.CO2.Hplus.one2four <- 10^-samdat$High.CO2.pH.one2four

row.names(samdat) = samdat$Sample.ID

# Create phyloseq object
ps = phyloseq(otu_table=otutab,sample_data=samdat)
ps = subset_samples(ps, Sample.ID != "049-W3-Compost" & Sample.ID!="004-K2-Muck" & Sample.ID!="028-S-0-30" & Sample.ID!="029-S-30-60")
```


```{r, results="hide"}
ps.melt = psmelt(ps)
ps.melt

head(ps.melt)

hist(sample_sums(ps))

# Relative abundances
ps.norm = transform_sample_counts(ps,function(x) x/sum(x))
head(otu_table(ps.norm))
sample_sums(ps.norm)
# Looks good

colnames(sample_data(ps.norm))
```

## Acidity Values

### Diversity Plots

```{r nmds-bray-soilph-Lab.CO2.pH.one2one, fig.height=4, fig.width=9}
ps.ord.nmds = ordinate(ps.norm, "NMDS", "bray",trymax=1000,k=2)
p = plot_ordination(ps.norm, ps.ord.nmds, type="samples")
palette = brewer.pal(8, "Spectral")
p = p + aes(colour=Lab.CO2.pH.one2one) + geom_point(size=3) + 
  scale_colour_gradientn(colors=palette) + facet_wrap(~Study) + 
  theme_bw()
p
```

```{r nmds-bray-soilph-Lab.CO2.pH.one2two, fig.height=4, fig.width=9}
ps.ord.nmds = ordinate(ps.norm, "NMDS", "bray",trymax=1000,k=2)
p = plot_ordination(ps.norm, ps.ord.nmds, type="samples")
palette = brewer.pal(8, "Spectral")
p = p + aes(colour=Lab.CO2.pH.one2two) + geom_point(size=3) + 
  scale_colour_gradientn(colors=palette) + facet_wrap(~Study) + 
  theme_bw()
p
```

```{r nmds-bray-soilph-Lab.CO2.pH.one2thre, fig.height=4, fig.width=9}
ps.ord.nmds = ordinate(ps.norm, "NMDS", "bray",trymax=1000,k=2)
p = plot_ordination(ps.norm, ps.ord.nmds, type="samples")
palette = brewer.pal(8, "Spectral")
p = p + aes(colour=Lab.CO2.pH.one2three) + geom_point(size=3) + 
  scale_colour_gradientn(colors=palette) + facet_wrap(~Study) + 
  theme_bw()
p
```

```{r nmds-bray-soilph-Lab.CO2.pH.one2four, fig.height=4, fig.width=9}
ps.ord.nmds = ordinate(ps.norm, "NMDS", "bray",trymax=1000,k=2)
p = plot_ordination(ps.norm, ps.ord.nmds, type="samples")
palette = brewer.pal(8, "Spectral")
p = p + aes(colour=Lab.CO2.pH.one2four) + geom_point(size=3) + 
  scale_colour_gradientn(colors=palette) + facet_wrap(~Study) + 
  theme_bw()
p
```

```{r richness-ph-Lab.CO2.pH.one2one, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.pH.one2one", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2two, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.pH.one2two", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2three, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.pH.one2three", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2four, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.pH.one2four", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2one, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.Hplus.one2one", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2two, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.Hplus.one2two", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2three, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.Hplus.one2three", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2four, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="Lab.CO2.Hplus.one2four", measures=c("Observed","Simpson","Shannon"), shape="Study")
p + geom_point(size=6, alpha=0.2) + theme_bw()
```















# Spooner Vs. Wisconsin

```{r subset-wisc-and-spooner}
ps.wisc <- subset_samples(ps, Study=="Wisconsin")
ps.spoon <- subset_samples(ps, Study=="Spooner")
```

```{r richness-ph-Lab.CO2.pH.one2one.wisc, fig.width=13, fig.height=4}
p.Lab.CO2.pH.one2one.wisc <- plot_richness(ps.wisc, x="Lab.CO2.pH.one2one", measures=c("Observed","Simpson","Shannon"))
p.Lab.CO2.pH.one2one.wisc + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2two.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="Lab.CO2.pH.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2three.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="Lab.CO2.pH.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2four.wisc, fig.width=13, fig.height=4}
p.Lab.CO2.pH.one2four.wisc <- plot_richness(ps.wisc, x="Lab.CO2.pH.one2four", measures=c("Observed","Simpson","Shannon"))
p.Lab.CO2.pH.one2four.wisc + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2one.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="Lab.CO2.Hplus.one2one", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2two.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="Lab.CO2.Hplus.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2three.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="Lab.CO2.Hplus.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2four.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="Lab.CO2.Hplus.one2four", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```


```{r richness-ph-Lab.CO2.pH.one2one.spoon, fig.width=13, fig.height=4}
p.Lab.CO2.pH.one2one.spoon <- plot_richness(ps.spoon, x="Lab.CO2.pH.one2one", measures=c("Observed","Simpson","Shannon"))
p.Lab.CO2.pH.one2one.spoon + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2two.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="Lab.CO2.pH.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2three.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="Lab.CO2.pH.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.pH.one2four.spoon, fig.width=13, fig.height=4}
p.Lab.CO2.pH.one2four.spoon <- plot_richness(ps.spoon, x="Lab.CO2.pH.one2four", measures=c("Observed","Simpson","Shannon"))
p.Lab.CO2.pH.one2four.spoon + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2one.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="Lab.CO2.Hplus.one2one", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2two.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="Lab.CO2.Hplus.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2three.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="Lab.CO2.Hplus.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-Lab.CO2.Hplus.one2four.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="Lab.CO2.Hplus.one2four", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```


```{r richness-ph-High.CO2.pH.one2one.wisc, fig.width=13, fig.height=4}
p.High.CO2.pH.one2one.wisc <- plot_richness(ps.wisc, x="High.CO2.pH.one2one", measures=c("Observed","Simpson","Shannon"))
p.High.CO2.pH.one2one.wisc + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.pH.one2two.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="High.CO2.pH.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.pH.one2three.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="High.CO2.pH.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.pH.one2four.wisc, fig.width=13, fig.height=4}
p.High.CO2.pH.one2four.wisc <- plot_richness(ps.wisc, x="High.CO2.pH.one2four", measures=c("Observed","Simpson","Shannon"))
p.High.CO2.pH.one2four.wisc + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2one.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="High.CO2.Hplus.one2one", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2two.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="High.CO2.Hplus.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2three.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="High.CO2.Hplus.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2four.wisc, fig.width=13, fig.height=4}
p <- plot_richness(ps.wisc, x="High.CO2.Hplus.one2four", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.pH.one2one.spoon, fig.width=13, fig.height=4}
p.High.CO2.pH.one2one.spoon <- plot_richness(ps.spoon, x="High.CO2.pH.one2one", measures=c("Observed","Simpson","Shannon"))
p.High.CO2.pH.one2one.spoon + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.pH.one2two.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="High.CO2.pH.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.pH.one2three.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="High.CO2.pH.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.pH.one2four.spoon, fig.width=13, fig.height=4}
p.High.CO2.pH.one2four.spoon <- plot_richness(ps.spoon, x="High.CO2.pH.one2four", measures=c("Observed","Simpson","Shannon"))
p.High.CO2.pH.one2four.spoon + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2one.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="High.CO2.Hplus.one2one", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2two.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="High.CO2.Hplus.one2two", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2three, fig.width=13, fig.height=4}
p <- plot_richness(ps, x="High.CO2.Hplus.one2three", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```

```{r richness-ph-High.CO2.Hplus.one2four.spoon, fig.width=13, fig.height=4}
p <- plot_richness(ps.spoon, x="High.CO2.Hplus.one2four", measures=c("Observed","Simpson","Shannon"))
p + geom_point(size=6, alpha=0.2) + theme_bw()
```



```{r diversity-plot}
plot_grid(p.Lab.CO2.pH.one2one.wisc, p.Lab.CO2.pH.one2four.wisc, p.High.CO2.pH.one2one.wisc, p.High.CO2.pH.one2four.wisc, p.Lab.CO2.pH.one2one.spoon, p.Lab.CO2.pH.one2four.spoon, p.High.CO2.pH.one2one.spoon, p.High.CO2.pH.one2four.spoon)
```







### Permanova Plot

```{r permanova-all-factors-ph}
df = as(sample_data(ps.norm), "data.frame")
d = phyloseq::distance(ps.norm, "bray") 
ps.adonis = adonis(d ~ Lab.CO2.pH.one2one+Lab.CO2.pH.one2two+Lab.CO2.pH.one2two+Lab.CO2.pH.one2three+Lab.CO2.pH.one2four+High.CO2.pH.one2one+High.CO2.pH.one2two+High.CO2.pH.one2two+High.CO2.pH.one2three+High.CO2.pH.one2four+Lab.CO2.Hplus.one2one+Lab.CO2.Hplus.one2two+Lab.CO2.Hplus.one2two+Lab.CO2.Hplus.one2three+Lab.CO2.Hplus.one2four+High.CO2.Hplus.one2one+High.CO2.Hplus.one2two+High.CO2.Hplus.one2two+High.CO2.Hplus.one2three+High.CO2.Hplus.one2four+DNA.Extr.pH.After.C1+DNA.Extr.pH.After.C2+DNA.Extr.Hplus.After.C1+DNA.Extr.Hplus.After.C2+Study, df)
ps.adonis
```

```{r permanova-all-treatments-ph}
df = as(sample_data(ps.norm), "data.frame")
d = phyloseq::distance(ps.norm, "bray") 
ps.adonis = adonis(d ~ Lab.CO2.pH.one2one+Lab.CO2.pH.one2two+Lab.CO2.pH.one2two+Lab.CO2.pH.one2three+Lab.CO2.pH.one2four+High.CO2.pH.one2one+High.CO2.pH.one2two+High.CO2.pH.one2two+High.CO2.pH.one2three+High.CO2.pH.one2four+DNA.Extr.pH.After.C1+DNA.Extr.pH.After.C2, df)
ps.adonis
```

```{r rsq-permanova-plot-ph}
rows.permanova <- rownames(ps.adonis$aov.tab)[1:10]
rows.permanova.nice <- c("1-to-1 Soil pH at 400ppm CO2","1-to-2 Soil pH at 400ppm CO2","1-to-3 Soil pH at 400ppm CO2","1-to-4 Soil pH at 400ppm CO2","1-to-1 Soil pH at 2% CO2","1-to-2 Soil pH at 2% CO2","1-to-3 Soil pH at 2% CO2","1-to-4 Soil pH at 2% CO2","Lysate pH in Buffer C1","Lysate pH in Buffer C2")
rsp.permanova <- ps.adonis$aov.tab[,5][1:10]
Suspension <- c("1-to-1","1-to-2","1-to-3","1-to-4","1-to-1","1-to-2","1-to-3","1-to-4","Buffer C1", "Buffer C2")
CO2.Level <- c("400ppm","400ppm","400ppm","400ppm","2%","2%","2%","2%","Laboratory","Laboratory")
dat.rsp.permanova <- data.frame(rows.permanova, rows.permanova.nice, rsp.permanova, Suspension, CO2.Level)
p <- ggplot(dat.rsp.permanova, aes(rsp.permanova, rows.permanova.nice, color=CO2.Level, shape=Suspension))
p + geom_point(size = 4) + theme_bw() + labs(y = "Treatment") +
  scale_x_continuous(name="PERMANOVA R-squared Value", limits=c(0, 0.075), breaks=c(0, .01, .02, .03, .04, .05, .06, 0.07))
```

```{r permanova-all-treatments-hplus}
df = as(sample_data(ps.norm), "data.frame")
d = phyloseq::distance(ps.norm, "bray") 
ps.adonis.hplus = adonis(d ~ Lab.CO2.Hplus.one2one+Lab.CO2.Hplus.one2two+Lab.CO2.Hplus.one2two+Lab.CO2.Hplus.one2three+Lab.CO2.Hplus.one2four+High.CO2.Hplus.one2one+High.CO2.Hplus.one2two+High.CO2.Hplus.one2two+High.CO2.Hplus.one2three+High.CO2.Hplus.one2four+DNA.Extr.Hplus.After.C1+DNA.Extr.Hplus.After.C2, df)
ps.adonis.hplus
```

```{r rsq-permanova-plot-hplus}
rows.permanova.hplus <- rownames(ps.adonis.hplus$aov.tab)[1:10]
rows.permanova.nice.hplus <- c("1-to-1 Soil a(H+) at 400ppm CO2","1-to-2 Soil a(H+) at 400ppm CO2","1-to-3 Soil a(H+) at 400ppm CO2","1-to-4 Soil Hplus at 400ppm CO2","1-to-1 Soil a(H+) at 2% CO2","1-to-2 Soil a(H+) at 2% CO2","1-to-3 Soil a(H+) at 2% CO2","1-to-4 Soil a(H+) at 2% CO2","Lysate a(H+) in Buffer C1","Lysate a(H+) in Buffer C2")
rsp.permanova.hplus <- ps.adonis.hplus$aov.tab[,5][1:10]
Suspension <- c("1-to-1","1-to-2","1-to-3","1-to-4","1-to-1","1-to-2","1-to-3","1-to-4","Buffer C1", "Buffer C2")
CO2.Level <- c("400ppm","400ppm","400ppm","400ppm","2%","2%","2%","2%","Laboratory","Laboratory")
dat.rsp.permanova.hplus <- data.frame(rows.permanova.hplus, rows.permanova.nice.hplus, rsp.permanova.hplus, Suspension, CO2.Level)
p <- ggplot(dat.rsp.permanova.hplus, aes(rsp.permanova.hplus, rows.permanova.nice.hplus, color=CO2.Level, shape=Suspension))
p + geom_point(size = 4) + theme_bw() + labs(y = "Treatment") +
  scale_x_continuous(name="PERMANOVA R-squared Value")
```





---

END
